Global Climate and Atmospheric Composition of the Ultra-Hot Jupiter WASP-103\lowercase{b} from \HST\ and \Spitzer\ Phase Curve Observations

### 2018-04-03 ApJ_submit_v2/response.txt

1       x
2       x
3       x
4       x
5	x
6       x
7       x
8
9       x
10	x
11      x
12
13
14      x-abstract
15
16      x
17      x
18
19	x
20
21
22      x
23

@@

I also realize that the content of this section may change depending on the previous comments regarding the interpretation of the 4.5 micron "emission feature". I think regardless of the outcome, it should be made clear how critically the metallicity figure that is obtained depends on accurately determining the strength of the emission at 4.5 microns and that if it is due to systematics that this analysis no longer applies.

****

14. The authors close section 5.2.1 with a list of caveats about the error bars on this section and how the inferred [M/H] and [C/O] may be heavily skewed by the one 4.5 micron data point (from phase 0.5? or other bin? see previous comment). These caveats need to be more prominent considering how serious they may be to the interpretation. I think they need to be present in the abstract, and also they need to be bulleted here in the relevant section so that they are easy to catch for those that skim the paper instead of read it in depth.

We have reformatted section 5.2.1 to include the caveats as a bulleted list. FIXME abstract

@@

15. Section 5.2.2: Nightside spectrum. The authors fit a nightside spectrum at phase 0.1. One easy thing to do for comparison and to estimate or show error bars is to also fit a nightside spectrum at phase 0.9, which should be symmetric and obtain the same or similar results. I would ask that the authors please do this and report the results. I realize it might not be interesting because the fit already presented is not able to say much about the metallicity or the C/O ratio. This may also be a useful exercise for the dayside spectrum (fit at phase 0.4 and phase 0.6, or something similar).

****

16. Section 6: The GCM used assumes a solar composition, but the authors argued earlier in the paper that WASP-103b's atmosphere is enhanced in metallicity. So why make this choice? Does the GCM not allow you to dial up the metallicity? This isn't explained. Later the authors also run an enhanced metallicity ([M/H] = 0.5) simulation with different drag coefficients. However, the reasons behind these different model choices for comparisons is never explained. There should be more text towards this end, particularly on the choice on whether or not to include TiO/VO opacity, which is turned on or off in the different simulations without explanation.

Each GCM run is very computationally intensive, so we ran a few representative cases to see which factors (TiO/VO, metallicity, drag) had the largest effect on the phase curve shape. Our nominal GCM was a cloud-free, solar composition model with TiO/VO and no added drag. We changed the parameters one at a time to isolate what their effects were (removing TiO/VO, increasing metallicity, including drag). We found that including drag, rather than changing metallicity, is what makes the biggest difference in the phase curve shape and provides the best match to the data (see Figure 15). We clarified these points in the text in section 6.

@@

****

17. The discussion of the magnetic field strength needed to produce a drag that sort of makes the GCM model work the best is very wishy washy. I understand that the authors are using order-of-magnitude expressions (as they state) but I think the magnetic field discussion is oversimplified and the conclusions are overstated. It also neglects to mention previous work attempting to directly measure/detect the effects of magnetic fields from planets. If the authors wish to discuss the magnetic fields, some mention of these previous works should be done. Significant mention about the shape of the magnetic field should also be prevalent. Furthermore, the planet is very close in to its star, and so there would likely be a complicated magnetic field structure due to the presence of the stellar magnetic field in the vicinity. This is not addressed or touched on when the authors assume the magnetic field is perpendicular to the flow of material.

If the authors want to keep a discussion of the magnetic field in the paper, this section should be significantly expanded and the caveats and assumptions more prominently listed. Otherwise, casual readers might over interpret this section to imply that a magnetic field is detected or that a magnetic field with a strength nice and similar to Jupiter's is the only possible explanation for the observable properties of the system with regards to its phase curve. It is the referee's preference however, that the magnetic field discussion is mostly scaled down or eliminated.

We expanded the discussion of magnetic fields according to the referee's suggestions. We think it is important to keep this section of the paper because even though this is just an order-of-magnitude result,

we want to make the point that the planet's magnetic field may make an observable difference to the atmospheric dyanamics, providing a novel way to study.

### 2018-04-03 ApJ_submit_v2/ms.tex
point 17

@@

We note that an important caveat for our results is that the best fit model is not a perfect fit to the data (with $\chi_\nu = 1.77$ for 9 degrees of freedom), so the uncertainties produced by the MCMC may be underestimated. Furthermore, the inferred C/O and metallicity are highly sensitive to the planet-to-star flux at \Spitzer\ $4.5\,\mu$m, which is the worst fit data point. To fit this data point, the model favors super-solar metallicities and C/O, which drive up the CO abundance (the dominant absorber at $4.5\,\mu$m). Given that the result hinges on a single photometric band and there are no spectrally resolved molecular features to constrain the temperature-pressure profile or which absorbers are present, we caution against over-interpreting these results until wider spectral coverage is available.

@@

Our nominal GCM was a cloud-free, solar composition atmosphere with TiO/VO opacity and no added drag.  We also ran models with enhanced metallicity ([M/H] = 0.5), no TiO/VO, and added atmospheric drag with timescales $\tau_\mathrm{drag} = 10^3$ and $10^4$ s, which we label $\tau_\mathrm{drag3}$ and $\tau_\mathrm{drag4}$, respectively.  To assess how well the GCM predictions reproduce the data, we calculated the amplitude and hot spot offset for all the models (listed in Table\,\ref{table:amps_offsets}). The small observed hotspot offsets ($-0.3 - 2.0$ degrees) are best reproduced by the $\tau_\mathrm{drag4}$ model, which has a smaller offset than the drag-free GCMs due to changes in wind pattern. In the drag models, the winds shift from a substellar to an antistellar flow rather than an equatorial jet.  The $\tau_\mathrm{drag4}$ model also provides the match the observed phase curve amplitudes.

@@

To estimate the magnetic field required to produce a fast drag timescale of order $10^4$\,s, we computed the free electron abundance due to ionized metals for the $\tau_\mathrm{drag4}$ GCM. Using the order-of-magnitude expression from \cite{perna10} (Equation 12), we estimate a drag timescale at the substellar point of $\tau_\mathrm{drag}  = 2\times10^4/B^2$ s, where $B$ is the magnetic field in Gauss. We assume a temperature and pressure of 3359 K and 0.11 bar, %(FIXME use correct values for drag4) 
and that the magnetic field is perpendicular to the flow. To reach a drag timescale of $10^4$\,s, a magnetic field stronger than $\sim1$ Gauss is required. This value is comparable to Jupiter's magnetic field strength of 5-10 Gauss \citep{bagenal04}.

We also compared the TP profiles from the $\tau_\mathrm{drag4}$ GCM to cloud condensation curves and the best fit radiative-convective equilibrium models from the 1D retrieval (Figure\,\ref{fig:TP}).  For the dayside photosphere, the TP profile slope and absolute temperature are in rough agreement between the 1D best fit and the GCM. At higher pressures, the GCM is systematically cooler, which is likely due to the effect of atmospheric circulation (at these pressures the GCM mixes the temperature planet-wide). At lower pressures, the GCM is also cooler than the 1D fits, which may be due to metallicity differences between the models. The GCM has solar metallicity, whereas the best fit 1D model has $[M/H] \sim 1$. Higher metallicity compositions have larger TiO/H2O ratios, and since the pressure dependence of TiO dissociation is not as strong as for water dissociation, we expect stronger inversions for higher metallicity atmospheres.  On the nightside, we also find that the GCM is cooler than the 1D models.  There are several physical processes not included in the GCM that could contribute to a hotter nightside, including shocks, longitude-dependent drag, and latent heat released from H2 recombination.  The best fit nightside TP profile is hotter than the condensation curves through most of the photosphere, suggesting that the observable atmosphere is relatively free of clouds. This prediction could be tested with longer wavelength phase curve observations.
%The GCM predicts cloud formation on the nightside, which lowers the temperature of the photosphere.

@@

>>>>

\newenvironment{my_itemize}{
    \begin{itemize}
              \setlength{\itemsep}{1pt}
                    \setlength{\parskip}{0pt}
                      \setlength{\parsep}{0pt}}{\end{itemize}
                      }

@@

\added{There are several caveats in our analysis which are important to note:}
\begin{my_itemize}
    \item {The best fit model is not a perfect fit to the data (with $\chi_\nu = 1.77$ for 9 degrees of freedom), so the uncertainties produced by the MCMC may be underestimated.} 
    \item {The inferred C/O and metallicity are highly sensitive to the planet-to-star flux at \Spitzer\ $4.5\,\mu$m, which is the worst fit data point. To fit this data point, the model favors super-solar metallicities and C/O, which drive up the CO abundance (the dominant absorber at $4.5\,\mu$m).}
    \item{\added{The \Spitzer\ $4.5\,\mu$m data is from broadband photometry, so the inferred CO feature is not spectrally resolved. It is possible that unknown absorbers or disequilibrium chemistry affect the $4.5\,\mu$m planet-to-star flux, but we cannot uniquely identify these features in our spectrum.}}
\end{my_itemize}
We therefore caution against over-interpreting these results until wider spectral coverage is available.

@@

Our nominal GCM was a cloud-free, solar composition atmosphere with TiO/VO opacity and no added drag.  \added{Each GCM run is computationally intensive, so we ran a small number of additional models to see which parameters had the largest effect on the planet spectrum. We changed model parameters one at a time,} considering cases with enhanced metallicity ([M/H] = 0.5), no TiO/VO, and added atmospheric drag with timescales $\tau_\mathrm{drag} = 10^3$ and $10^4$ s, which we label $\tau_\mathrm{drag3}$ and $\tau_\mathrm{drag4}$, respectively.  To assess how well the GCM predictions reproduce the data, we calculated the amplitude and hot spot offset for all the models (listed in Table\,\ref{table:amps_offsets}). The small observed hotspot offsets ($-0.3 - 2.0$ degrees) are best reproduced by the $\tau_\mathrm{drag4}$ model, which has a smaller offset than the drag-free GCMs due to changes in wind pattern. In the drag models, the winds shift from a substellar to an antistellar flow rather than an equatorial jet.  The $\tau_\mathrm{drag4}$ model also provides the match the observed phase curve amplitudes.

@@

We also compared the TP profiles from the $\tau_\mathrm{drag4}$ GCM to cloud condensation curves and the best fit radiative-convective equilibrium models from the 1D retrieval (Figure\,\ref{fig:TP}).  For the dayside photosphere, the TP profile slope and absolute temperature are in rough agreement between the 1D best fit and the GCM. At higher pressures, the GCM is systematically cooler, which is likely due to the effect of atmospheric circulation (at these pressures the GCM mixes the temperature planet-wide). At lower pressures, the GCM is also cooler than the 1D fits, which may be due to metallicity differences between the models. The GCM has solar metallicity, whereas the best fit 1D model has $[M/H] \sim 1$. Higher metallicity compositions have larger TiO/H$_2$O ratios, and since the pressure dependence of TiO dissociation is not as strong as for water dissociation, we expect stronger inversions for higher metallicity atmospheres.  On the nightside, we also find that the GCM is cooler than the 1D models.  There are several physical processes not included in the GCM that could contribute to a hotter nightside, including shocks, longitude-dependent drag, and latent heat released from H2 recombination.  The best fit nightside TP profile is hotter than the condensation curves through most of the photosphere, suggesting that the observable atmosphere is relatively free of clouds. This prediction could be tested with longer wavelength phase curve observations.
%The GCM predicts cloud formation on the nightside, which lowers the temperature of the photosphere.

@@

\subsection{Constraints on the Planet's Magnetic Field}
\added{We showed in this section that the small observed hotspot offset in the phase curves is best fit by a GCM that includes magnetic drag with a timescale $\tau_\mathrm{drag} = 10^4$\,s. This observation gives rise to the question: what magnetic field strength on the planet can produce drag with this timescale? Previous efforts to characterize exoplanet magnetic fields have mainly focused on magnetic interaction between the planet and its host star \citep[e.g.][and references therein]{wright17}.  A complementary approach is to study the effect of the magnetic field on the planet's atmospheric dynamics.}

Here we make a simple order-of-magnitude estimate for the magnetic field required to produce a drag timescale of order $10^4$\,s. We first computed the free electron abundance due to ionized metals for the $\tau_\mathrm{drag4}$ GCM. Using the analytic expression from \cite{perna10} (Equation 12), we estimated a drag timescale at the substellar point of $\tau_\mathrm{drag}  = 2\times10^4/B^2$ s, where $B$ is the magnetic field in Gauss. We assumed a temperature and pressure of 3359 K and 0.11 bar, %(FIXME use correct values for drag4) 
and that the magnetic field is perpendicular to the flow. To reach a drag timescale of $10^4$\,s, a magnetic field stronger than $\sim1$ Gauss is required, comparable to Jupiter's magnetic field strength of 5-10 Gauss \citep{bagenal04}.  \added{To confirm this intriguing result, more detailed study is warranted, including a full magnetic hydrodynamic simulation of the atmospheric dynamics \citep[e.g.][]{rogers17} and accounting for possible complex magnetic field structure due to interactions between the magnetic fields of the planet and the star.}
